\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors

### 2022-06-15 Manuscript/Lanuza_et_al_2021.Rmd
edits

Although the negative correlation between flower number and flower size has received substantial attention [worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. Here, we provide support to this evidence with a great diversity of phylogenetic lineages (170 families and over 1000 species) and show that other reproductive and plant size and form related traits also covariate along this axis. For example, most species with unisexual flowers (14% of all species) were concentrated in the trait space and were associated with species with woddy life forms (i.e., trees and shrubs) with high flower number. This empirical evidence corroborates the idea that approximately 10% of species have separation of sexes [@barrett2002] and that breeding systems are associated with specific life forms [@petit2006]. The pollinator dependence - floral display axis

check number of species of goodwillie and what we offer that is new...

Further, the pollinator dependence - floral display axis provided additional support for the previously described positive association between higher outcrossing rate and larger floral display [@goodwillie2010].

Style length...

@@

\vspace{-28mm}

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=12, fig.width=18, out.width="90%",fig.align='center',fig.cap= "\\textbf{Figure 2} Location of the different qualitative traits on the trait space. The panel is composed by the traits that showed statistical association with the first two axes of trait variation: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f). The solid arrows indicate the direction of variation of the different quantitative traits showed in figure 1: flower number (FN), plant height (PH), style length (SL), flower size (FS), ovule number (ON) and the level of autonomous selfing (S)."}

>>>>

Our study provides solid evidence for several widely discussed trait associations with the help of a great diversity of phylogenetic lineages (170 families and over 1000 species) within a multitrait context. Although the negative correlation between flower number and flower size has received substantial attention [worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. We found support to this trait association and show that other reproductive and plant size and form related traits also covariate along this axis. For example, our study complements previous evidence that links breeding systems and life forms [@petit2006; @friedman2020] by showing that most species with unisexual flowers were concentrated in the trait space with woody life forms with high flower number. Moreover, our work also confirms the previously described positive association between outcrossing rate and larger floral display [@goodwillie2010] that has been reviewed several times as one of the most common mechanisms to promote pollen transfer and guarantee reproductive success.

Our work also confirms the previously described positive association between outcrossing rate and floral display

Plants with larger floral displays can ensure pollen transfer

A common strategy of flowering plants to ensure pollen transfer by allocating more resources to floral display

Flowering plants have developed multiple strategies to influence pollinator behaviour and ensure pollen transfer [@minnaar2019]. One of the most well known trait associations is the greater allocation of resources from oucrossing species to floral display

Plants in order to ensure reproductive success have developed a broad range of reproductive strategies.

Because most flowering plants rely on pollinators to ensure reproductive success [@ollerton2011], plants have developed multiple strategies to influence pollinator behaviour and ensure pollen transfer [@minnaar2019]. We found support to the idea that

Resource allocation to attraction has been shown to influence pollinator visitation and impact reproductive success

Our work also confirms the previously described positive association between outcrossing rate and floral display [@goodwillie2010] where species with larger floral displays

Similarly, the second axis of reproductive trait variation also confirmed the previously described positive association between outcrossing rate and floral display [@goodwillie2010]. Resource allocation to attraction has been shown to influence pollinator visitation and impact reproductive success [@minnaar2019]

Plants with larger floral displays have been shown to attract more floral visitors and ensure reproductive success

that is thought to be one of the mechanisms to ensure pollen transfer [@minnaar2019].

As expected, species with higher pollinator dependence showed greater allocation of resources to attraction that has been  in order to ensure pollen transfer

species with higher pollination dependende had greater allocation to flower number or flower size.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=12, fig.width=18, out.width="80%",fig.align='center',fig.cap= "\\textbf{Figure 2} Location of the different qualitative traits on the trait space. The panel is composed by the traits that showed statistical association with the first two axes of trait variation: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f). The solid arrows indicate the direction of variation of the different quantitative traits showed in figure 1: flower number (FN), plant height (PH), style length (SL), flower size (FS), ovule number (ON) and the level of autonomous selfing (S)."}

### 2022-06-15 Manuscript/Lanuza_et_al_2021.tex
edits

Although the negative correlation between flower number and flower size has received substantial attention {[}worley2000; \protect\hyperlink{ref-sargent2007}{Sargent et al.} (\protect\hyperlink{ref-sargent2007}{2007}); \protect\hyperlink{ref-kettle2011}{Kettle et al.} (\protect\hyperlink{ref-kettle2011}{2011}){]}, it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. Here, we provide support to this evidence with a great diversity of phylogenetic lineages (170 families and over 1000 species) and show that other reproductive and plant size and form related traits also covariate along this axis. For instance, most unisexual species (14\% of plant species of our dataset) were associated with woody species with high flower number. This concurs well with the previous idea that approximately 10\% of flowering plants have separation of sexes (\protect\hyperlink{ref-barrett2002}{Barrett, 2002}) and that species with unisexual flowers tend to be associated with woody/tree life forms (\protect\hyperlink{ref-petit2006}{Petit \& Hampe, 2006}).

The second axis organized the different species by the degree of pollinator dependence where species with higher pollinator dependence showed greater floral display.

As expected\ldots{} quizas hilar aqui con un par de refs para que quede menos results

@@

@@

\vspace{-28mm}

@@

{\centering \includegraphics[width=0.9\linewidth]{output/figures/unnamed-chunk-3-1}

>>>>

Our study provides solid evidence for several widely discussed trait associations with the help of a great diversity of phylogenetic lineages (170 families and over 1000 species) within a multitrait context. Although the negative correlation between flower number and flower size has received substantial attention {[}worley2000; \protect\hyperlink{ref-sargent2007}{Sargent et al.} (\protect\hyperlink{ref-sargent2007}{2007}); \protect\hyperlink{ref-kettle2011}{Kettle et al.} (\protect\hyperlink{ref-kettle2011}{2011}){]}, it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. We found support to this trait association and show that other reproductive and plant size and form related traits also covariate along this axis. For example, our study complements previous evidence that links breeding systems and life forms {[}\protect\hyperlink{ref-petit2006}{Petit \& Hampe} (\protect\hyperlink{ref-petit2006}{2006}); friedman2020{]} by showing that most species with unisexual flowers were concentrated in the trait space with woody life forms with high flower number. Similarly, the second axis of reproductive trait variation also confirmed the previously described positive association between outcrossing rate and floral display (\protect\hyperlink{ref-goodwillie2010}{Goodwillie et al., 2010}).

that is thought to be one of the mechanisms to ensure pollen transfer (\protect\hyperlink{ref-minnaar2019}{Minnaar et al., 2019}).

As expected, species with higher pollinator dependence showed greater allocation of resources to attraction that has been in order to ensure pollen transfer

species with higher pollination dependende had greater allocation to flower number or flower size.

@@

\leavevmode\vadjust pre{\hypertarget{ref-minnaar2019}{}}%
Minnaar, C., Anderson, B., Jager, M. L. de, \& Karron, J. D. (2019). Plant--pollinator interactions along the pathway to paternity. \emph{Annals of Botany}, \emph{123}(2), 225--245.

@@

@@

{\centering \includegraphics[width=0.8\linewidth]{output/figures/unnamed-chunk-3-1}
